### 2013-05-07 research proposal MTS.tex
Started with research proposal MTS

\documentclass[12pt]{article}

\title{Research proposal: idiolects in word prediction}
\author{Wessel Stoop}

\usepackage{graphicx}
\usepackage{float}
\usepackage{apacite}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}

\section{Project title}
Idiolects in word prediction
\section{Summary}
[[]]
\section{Candidate}
Wessel Stoop
\section{Description of the proposed research}

\section{Summary in keywords}

\end{document}